---
title: "assignment 5"
author: "vineeth goud maddi"
date: "4/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
View(Cereals)
library(cluster)
library(caret)
library(dendextend)
library(factoextra)

```
data importing cereals dataset
```{r}
data.frame <-data.frame(Cereals[,4:16])
```
data processing. removing the missing values that might present in the data
```{r}
removed_missingvalue <- na.omit(data.frame)
#Data normalization and data scaling
Normalize <- scale(removed_missingvalue)
```
using the euclidean distance to measure the distance
```{r}
d <- dist(Normalize, method = "euclidean")
#perform hierarchical clustering using complete linkage.
Hc <- hclust(d, method = "complete")
plot(Hc)
#dendogram
round(Hc$height, 3)
```
Determining Optimal clusters:
highliting the clusters in dendogram directly.
```{r}
plot(Hc)
rect.hclust(Hc,k = 4, border = "orange")
```
We can also use agnes() function to perform clustering.
Performing clustering using agnes() with single, complete, average and ward.
```{r}
Hcsingle  <- agnes(Normalize, method = "single")
Hccomplete <-agnes(Normalize, method = "complete")
Hcaverage <-agnes(Normalize, method = "average")
Hcward <- agnes(Normalize, method = "ward")
```
Compare the agglomerative coefficients for single,complete,average and ward.
```{r}
print(Hcsingle$ac)
print(Hccomplete$ac)
print(Hcaverage$ac)
print(Hcward$ac)
```
From the above results the best value we got is 0.904. Ploting the agnes using ward method and cuttung the Dendrogram. we will take k =4 by observing the distance
```{r}
pltree(Hcward, cex = 0.6, hand = -1, main = "Dendrogram of agnes ward")
```
Hierarchical clustering using ward method.
```{r}
hc1 <- hclust(d, method = "ward.D2")
subgroup <- cutree(hc1, k =4)
table(subgroup)
datafram <- as.data.frame(cbind(Normalize,subgroup))
#the results in scatter plot.
fviz_cluster(list(data = Normalize,cluster=subgroup))
```
```{r}
datacereals <-Cereals
datacereals.omi <- na.omit(datacereals)
clust <- cbind(datacereals.omi, subgroup)
clust[clust$subgroup==1,]
clust[clust$subgroup==2,]
clust[clust$subgroup==3,]
clust[clust$subgroup==4,]
```
calculating the mean ratings to determine the cluster cereals
```{r}
mean(clust[clust$subgroup==1,"rating"])
mean(clust[clust$subgroup==2,"rating"])
mean(clust[clust$subgroup==3,"rating"])
mean(clust[clust$subgroup==4,"rating"])
```
from the above results we can clearly that the mean rating is high for subgroup 1. 